Hybrid quantum-classical unsupervised data clustering based on the Self-Organizing Feature Map

### 2024-11-11 manuscript/main.tex
Expand tabs
Fix "No Society specified, using generic APS style"
See https://tex.stackexchange.com/a/19585
No type size specified, using default 10
Add draft of scalling discussion

\documentclass[pra,showkeys,twocolumn,showpacs]{revtex4-1}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{array}
\usepackage{color}
\usepackage{soul}
%\usepackage{floatflt}
\usepackage{graphicx}
\DeclareMathOperator{\tr}{Tr}
\newcommand*{\eq}[1]{(\ref{#1})}
\newcommand*{\eqs}[2]{(\ref{#1})--(\ref{#2})}
\newcommand{\exi}{\tr{\{e^{-\frac\beta2\sum_k\lambda_k\gamma_k^\dag\gamma_k}\}}}
\renewcommand{\l}{\left(}
\renewcommand{\r}{\right)}

@@

>>>>

\documentclass[pra,showkeys,twocolumn,showpacs,aps,10pt]{revtex4-1}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{array}
\usepackage{color}
\usepackage{soul}
%\usepackage{floatflt}
\usepackage{graphicx}
\DeclareMathOperator{\tr}{Tr}
\newcommand*{\eq}[1]{(\ref{#1})}
\newcommand*{\eqs}[2]{(\ref{#1})--(\ref{#2})}
\newcommand{\exi}{\tr{\{e^{-\frac\beta2\sum_k\lambda_k\gamma_k^\dag\gamma_k}\}}}
\renewcommand{\l}{\left(}
\renewcommand{\r}{\right)}

@@

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{tomography-errors-by-shots-n4.eps}
  \caption{%
  }
  \label{fig:tomography-errors-by-shots-n4}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{tomography-errors-by-shots-n5.eps}
  \caption{%
  }
  \label{fig:tomography-errors-by-shots-n5}
\end{figure}

\section{Scalability and robustness}

Most quantum algorithms face three main scaling problems.
The first is the exponential increase in the number of iterations of the quantum circuit required.
This is not necessary in QASOFM\@.
SOFM assumes that the cluster points do not have to move monotonically towards the cluster centroids,
and that the vectors closest to the true BMU have the highest probability of being measured.
The number of runs should only be chosen based on a given accuracy of BMU selection.
It will only affect the speed of reaching the centres of the clusters.
Fig.~\ref{fig:tomography-errors-by-shots-n4} and Fig.~\ref{fig:tomography-errors-by-shots-n5} show the dependence of the number of tomography errors of vectors
which differ by no more than one bit (see Fig.~\ref{fig:distance_matrix}) on the number of shots.
Here, tomography implies the determination of all pairs of nearest vectors in the whole state space.
In Fig.~\ref{fig:tomography-errors-by-shots-n4} shows the state space of $2^4$ vectors, and Fig.~\ref{fig:tomography-errors-by-shots-n5} --- $2^5$ vectors.
The points in Fig.~\ref{fig:tomography-errors-by-shots-n4} and Fig.~\ref{fig:tomography-errors-by-shots-n5} are the number of misidentified neighbors
by the quantum circuit run on a noise simulator taking into account parameters of the real quantum platform.
The lines are the result of a regression analysis of these data.
The coefficient of determination of the model is $R^2 = 0.82$ for the state space of $2^4$ and $R^2 = 0.9$ for $2^5$.
The triangles are the number of neighbors misidentified by the quantum scheme run on the IBM quantum backend named ``ibm\_sherbrook''.
The number of errors decreases exponentially as the number of shots increases linearly.
In contrast to QASOFM,
tomography with increasing state space requires an exponential increase in the number of shots
to distinguish neighbors of different orders.

The second problem is a linear increase in a number of connections required for a single qubit.
In QASOFM, this is the ancilla that accumulates distances in phase.
Currently, this is a significant limitation for scaling on platforms with a rigid topology,
such as IBM or Google Sycamore processors.
For this reason, we cannot implement an algorithm for a larger problem.
The only solution is to create a large auxiliary network,
but this requires numerous two-qubit operations,
which are currently performed with insufficient accuracy.
Other approaches to implementing the quantum platform,
for example for ions in traps or Bose-Einstein condensates,
are likely to overcome this limitation.

The third problem is the difficulty of converting digital data into a quantum state.
The dimension of the data is usually much larger than the number of cluster vectors.
Consequently, the size of the subspace stretched over the cluster vectors will be negligible compared to the whole space.
This makes it possible to efficiently generate quantum circuits to initialise such superposition quantum states.
